\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-07-11 papers/methods/paper.tex
Updating plots, paper and data

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

\begin{equation}
M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{other corrections}, \label{eq:standard}
\end{equation}

@@

\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C),
\end{equation}

@@

\begin{align}
\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right],\\
r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
%\mu_{\rm obs} &= m_B + \alpha x_1 - \beta c - M_B. \label{eq:mu1}
\end{align}
where $r$ is the comoving distance for redshift $z$ given a specific cosmology and $H_0$ is the current value of Hubble's constant in $\kmsmpc$.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{plot_pop_simple.pdf}
	\end{center}
	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys. Selection effects are visible in both samples, where red supernovae are often cut as redshift increases. The colour of the data points is representative over the supernovae colour itself, blue dots represent blue supernovae, with warmer colours representing redder supernovae.}
	\label{fig:simple_pop}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
	\end{center}
	\caption{Posterior surfaces for 100 realisations of supernova data with the $\Lambda$CDM model. Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
	\label{fig:simple_ol}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
	\end{center}
	\caption{Posterior surfaces for 100 realisations of supernova data with the Flat $w$CDM model. The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the highly non-Gaussian nature of the posterior surface, where the point attained from each parameters independent maximum likelihood lands well outside the $3\sigma$ contour in parameter space.}
	\label{fig:simple_w}
\end{figure}

\begin{table}
	\centering
	\caption{Maximum likelihood cosmological parameters determined from stacking the surfaces of 100 fits to independent realisations of toy supernova data. As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae. Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$. As our surfaces are stacked, a shift away from the simulated values of more than $\sigma/\sqrt{100} = 0.1\sigma$ would give a $1\sigma$ detection of bias. No significant bias is detected in either cosmological model.}
	\label{tab:simple_model}
	\begin{tabular}{l|cc}
		\hline
		Model & $\Omega_m$ & $w$ \\ 
		\hline
		Flat $\Lambda$CDM & $0.302\pm 0.015$ & -- \\ 
		Flat $w$CDM + $\Omega_m$ prior & $\left( 300.9\pm7.4 \right) \times 10^{-3}$ & $-0.996\pm 0.038$ \\ 
		\hline
	\end{tabular}
\end{table}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{bulk_small_g10.pdf}		\includegraphics[width=\columnwidth]{bulk_small_c11.pdf}
	\end{center}
	\caption{Posterior surfaces for 100 realisations of supernova data for two intrinsic dispersion models - the {\gten} model for the top panel and the {\celeven} model for the bottom panel. Surfaces for $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$ are shown, with the other fit parameters being marginalised over. The intrinsic scatter model has significant impact on the recovered $\beta$, which in itself effects cosmology, resulting in small biases in $w$.}
	\label{fig:bulk_posterior}
\end{figure}

>>>>

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3\vspace{-5mm}
\end{keywords}

@@

\begin{align}
M_B = m_B + \alpha x_1 - \beta c - \mu(z) + \text{other corrections}, \label{eq:standard}
\end{align}

@@

\begin{align}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1} (\mu_{\rm obs} - \mu_C),
\end{align}

@@

\begin{align}
\mu_C &= 5 \log\left[ \frac{(1+z)r}{10} \right],\\
r& =\frac{c}{H_0} \int_0^z \frac{dz'}{\sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}}, \label{eq:mu0}
\end{align}
where $r$ is the comoving distance for redshift $z$ given a specific cosmology and $H_0$ is the current value of Hubble's constant in $\kmsmpc$.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{plot_pop_simple.pdf}
	\end{center}
	\caption{Population distributions shown in redshift and uncorrected absolute magnitude $m_B - \mu$ for 1000 supernovae in both high-redshift and low-redshift surveys. Selection effects are visible in both samples, where red supernovae are often cut as redshift increases. The colour of the data points is representative over the supernovae colour itself, blue dots represent blue supernovae, with warmer colours representing redder supernovae.}
	\label{fig:simple_pop}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelOl.pdf}
	\end{center}
	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $\Lambda$CDM model, with a representative contour from a single data realisation shown for context. Even a large supernova sample when treated robustly is insufficient to provide tight constraints on either $\Omega_m$ and $\Omega_\Lambda$ due to the severe degeneracy between the parameters.}
	\label{fig:simple_ol}
\end{figure}
\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{simpleApproximateModelW.pdf}
	\end{center}
	\caption{Maximal posterior points for 100 realisations of supernova data with the Flat $w$CDM model, with a representative contour from a single data realisation shown for context. The well known banana shaped contour is recovered, with the marginalised distributions in $\Omega_m$ and $w$ providing misleading statistics due to the highly non-Gaussian nature of the posterior surface. The recovered posteior maximums show the same degeneracy direction as the representative surface and scatter around the truth values input into the simulation, shown in dashed lines.}
	\label{fig:simple_w}
\end{figure}

\begin{table}
	\centering
	\caption{Cosmological parameters determined from the surfaces of 100 fits to independent realisations of toy supernova data. As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae. For each chain, we record the mean and standard deviation, and then show the average mean and average standard deviations in the table. The scatter introduced by simulation variance (the standard deviation of the 100 mean parameter values) is shown in brackets. Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$. As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted standard deviation: $\sqrt{100} = 10$. As the deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model (which was run with a prior on $\Omega_m$).}
	\label{tab:simple_model}
	\begin{tabular}{l|cc}
		\hline
		Model & $\Omega_m$(scatter) & $w$(scatter) \\ 
		\hline
		Flat $\Lambda$CDM & $0.301\pm 0.015(0.012)$ & -- \\ 
		Flat $w$CDM & -- & $-1.00\pm 0.042(0.030)$ \\ 
		\hline
	\end{tabular}
\end{table}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{bulk_points_g10.pdf}		\includegraphics[width=\columnwidth]{bulk_points_c11.pdf}
	\end{center}
	\caption{Posterior surfaces for 100 realisations of supernova data for two intrinsic dispersion models - the {\gten} model for the top panel and the {\celeven} model for the bottom panel. Surfaces for $\Omega_m$, $w$, $\alpha$, $\beta$ and $\langle M_B \rangle$ are shown, with the other fit parameters being marginalised over. The intrinsic scatter model has significant impact on the recovered $\beta$, which in itself effects cosmology, resulting in small biases in $w$.}
	\label{fig:bulk_posterior}
\end{figure}
